\SW: II. New lens candidates from the CFHTLS discovered through citizen science

### 2015-04-09 doc/sw-cfhtls.tex
rearranged results section and added a few para on citizens value; phil and i will make edits together on this version

We report the discovery of 61 new lens candidates in the CFHT Legacy Survey data
(\cfhtls) based on about 11 million classifications performed by citizen
scientists as part of the first blind lens search undertaken by \sw.  The
principle goal of this lens search was to identify lenses missed by lens finding
robots (the \rf on galaxy scales and \af on group/cluster scales), which have
been previously used to mine the \cfhtls data for lenses.  We compare the
properties of known lens samples detected by these algorithms to the \sw sample
and found that they are broadly similar.  The image separation distribution
calculated from the combined sample of known and new lens candidates shows that
our previous constraints on the average density profile of the lens population
are robust.  In addition to recovering about 60\% of the known sample, the new
\sw candidates possess a richer variety compared to the lenses found by the two
robots. We find that analyzing only those classifications which are performed by
the more skilled volunteers, \sw can achieve a detection rate of up to 75\%.
Training and calibration of the performance of citizen scientists is crucial for
the success of \sw. We also present the SIMCT pipeline, used for generating a
sample of realistic simulated lensed images in the CFHTLS, and a sample of duds
and false positives used in the training.  Such a training sample has a legacy
value for testing future lens finding algorithms. We make our training sample
publicly available.

%In summary, \sw
%has been successful in discovering a large sample of new lens candidates in
%addition to recovering a substantial fraction of the known lenses in this pilot
%lens search.

\end{abstract}
%The simulated sample is also used to make a comparison between human performance against the robots.

@@

The last few decades have seen a rise in the discoveries of strong
gravitational lenses owing to the plethora of interesting applications
lenses have in astrophysics and cosmology. Strong lenses are routinely
used to probe the dark matter distribution from galaxy (REF) to cluster
scales (REF), to  study distant young galaxies by using the lensing
magnification as a natural telescope (REF), to test the cosmological
model by constraining cosmological parameters such as the Hubble
constant (REF) and dark energy (REF) and many more. Strong
lenses are rare since a foreground massive object needs to be
sufficiently aligned with a distant background source to produce
multiple images. Nevertheless, systematic lens searches have led to the
discovery of over 500 lenses till
date\footnote{http://masterlens.astro.utah.edu}.

The search for gravitational lenses is a needle-in-a-haystack problem.
Efficient automated methods are imperative to find a reasonably complete and
pure sample of strong lenses. Several lens finding algorithms have been
developed so far for this purpose (REF) but they can not simultaneously capture
the myriad types of lenses that are known to exist.  For example, the lensed
images of background galaxies show variety in their surface brightness
distributions, colors, light profiles, shapes, structures and angular image
separations.  Moreover, many lensed images appear similar to features found
commonly in galaxies (such as spiral arms) or to artifacts in astronomical
images (scattered light around stars).  Almost all lens finding algorithms find
it difficult to distinguish these from the real lenses and thus, suffer from a
high rate of false positives.  To minimize this problem, algorithms are often
restricted to detect a very narrow class of lens systems. However, even after
such restrictions, robotic lens searches have to always rely on visual screening
to arrive at a list of plausible candidates.

Recognising patterns amidst an array of noisy pixelated images is one of the
strengths of human brains (REF). Humans are also capable of dealing with
multi-tier complex web of questions before arriving at a conclusion which may
not be always possible to automate. The algorithm by which our brains process a
task is extremely malleable, self-learning and self-evolving. Therefore it has
a huge potential for the discovery of exotic objects which do not quite fit a
set criteria, but are still very likely to be objects of interest. The lens
finding algorithms are not yet advanced enough to produce better performance
than visual classifications. Given this, participation of a large community of
volunteers to help with the visual identification of lenses is extremely
beneficial for the lensing community. In particular, as we enter the era of
large area imaging surveys (spanning thousands of square degrees), it is
the perfect time to tap into the potential of citizen science.

\GZ was the first citizen science project in astronomy, and addressed the
problem of how to  classify large numbers of galaxies by their morphology
\citep{Lintott2008}.  From the early results based on morphological
classifications of galaxies, to several new unexpected and interesting
discoveries, such as that of green pea galaxies and Hanny's Voorwerp, \GZ
has been able to harness the potential of citizen scientists.  Since then
astronomy and non-astronomy projects have been launched under the citizen
science web portal Zooniverse (http://zooniverse.org). The task of finding
gravitational lenses is significantly challenging given the fact that these
lenses are rare. To add to the challenge, not many citizen scientists are
expected to be aware of the phenomenon of gravitational lensing, and the
resulting image configurations.  With these significant challenges at hand, we
designed the \sw project to enable the discovery of lenses through citizen
science.  In Paper I (REF), we describe the design of \sw and how the entire
system functions as a discovery engine.

%
%\subsection{Paper I: Overview}
%In Paper I, we give details of the classification interface used by the
%volunteers. The regular images called test subjects are served randomly from the
%database. Some of the test subjects are labelled as training subjects because
%either these have simulated lensed images superposed around real galaxies or
%there are some false positives in the subjects.  The sample of training subjects
%is much smaller compared to the sample of test subjects. The training subjects
%are shown with decreasing frequency to the volunteers depending on how new the
%volunteers are to the project based on their classification history. However,
%the frequency of training subjects becomes a constant after XXX classifications
%from a volunteer.
%
%Clicking (or failing to click) on the training subjects results in a
%pop-up feedback message that congratulates the volunteers or points them
%to read the Spotter's guide pages. The website has several examples of
%real lenses and false positives with explanation on the Spotter's guide
%page. A quick reference to the Spotter's guide is available in the
%classification interface as well. Additional features to change zoom,
%contrast and brightness are made available to make a better judgement of
%the difficult lens candidates.
%
%The classification of subjects happens in two serial stages. \StageOne is
%for quick rejection of all uninteresting subjects. Subjects that passed
%\StageOne are then reinspected at \StageTwo, for a careful selection of
%most promising lens candidates. In general, subjects are assigned a
%small prior probability for it to contain a lens. As each subject is
%seen by multiple volunteers, the posterior probability P is updated by
%using the weights of the volunteers. The weights are determined based on
%how well the volunteers perform on the training sample. Once a subject
%crosses a threshold (acceptance or rejection), it is retired from the
%database to ensure that rest of the subjects receive sufficient multiple
%classifications. This is true for \StageOne with ~10$^5$ subjects but is
%not required for \StageTwo with only a few thousand subjects.
%
%{\bf AM: add one more paragraph that gives some key conclusions and numbers
%from Paper I}
%\question{AM}{PJM,AV, should we have this overview section on paper I?}
%

In this paper ( Paper II hereafter), we describe our first lens search using
data from the Canada-France-Hawaii Telescope Legacy Survey (\cfhtls, REF).  This
paper is organised as follows. In \Sref{sec:data}, we introduce the \cfhtls
imaging data and the known lens samples from the \cfhtls. In \Sref{sec:ts} and
\Sref{sec:dfp}, we explain the training sample consisting of simulated lenses,
duds and false positives generated to aid the \sw volunteers in the lens search.
In \Sref{sec:results}, we present the new lens candidates from \sw and compare
\sw to two lens finding algorithms which produced the known \cfhtls lens
samples. Next, we discuss what kind of lenses are detected or missed by the
algorithms and \sw in \Sref{sec:discuss}. The summary and conclusions are given
in \Sref{sec:conclude}.

@@

The \cfhtls is a photometric survey in five optical bands
($u^*g'r'i'z'$) carried out with the wide-field imager MegaPrime which
has a 1~deg$^2$ field-of-view and a pixel size of 0.186\arcsec. The
\cfhtls \textsc{WIDE} covers a total area of 171~deg$^2$ on the sky and
consists of four fields W1, W2, W3 and W4. The field W1 has the
largest sky coverage of 63.65~deg$^2$. The fields W2 and W4 have similar
sky coverages of 20.32~deg$^2$ and 20.02~deg$^2$,
respectively\footnote{These numbers are estimated from
http://terapix.iap.fr/cplt/table\_syn\_T0006.html}.  The field W3 has a
sky coverage of 42.87~deg$^2$ and is more than twice as large as W2 and
W4.

The \cfhtls imaging is very homogeneous and has great image quality. Most of
the lensed arcs are much brighter in the $g$ band thus, deep imaging in
this band is desirable. The limiting magnitude is 25.47 for the $g$ band
which goes the deepest among all of the five bands. The mean seeing in
the $g$ band is 0.78\arcsec. The zero point to convert flux to AB
magnitude for all bands is 30. These characteristics make \cfhtls ideal
to do visual inspection for finding lenses.  We use the data from the
final T0007 release taken from the Terapix
website\footnote{{http://terapix.iap.fr/cplt/T0006-doc.pdf}}
for this work.

@@

@@

In order to perform a blind lens search over the entire \cfhtls \textsc{wide},
we present the users with cutouts of images selected from the survey region. We
use the $g$, $r$ and $i$-band imaging from \cfhtls which are most useful for visual
identification of lenses. We also provided access to the $u$ and $z$ band imaging
data through dashboard, when users wished to take a closer look at a particular system\footnote{The dashboard tool page can be accessed from
http://tools.zooniverse.org/\#/dashboards/spacewarp}.

We made color composite images using the publicly-available
code HumVI\footnote{The open source  color image composition code used in this
work is available from \texttt{http://github.com/drphilmarshall/HumVI}}
following the prescription of \citet{Lupton2004}. The code
rescales the pixel values of each channel image into flux units, and
then applies an arcsinh stretch. The stretch parameters are chosen
using a small random sample of images, to ensure that the background
noise is just visible, and that the centres of bright, intermediate
redshift galaxies are not saturated. The color scales are chosen to
maximize the contrast between faint extended objects. These parameters
are then fixed during the production of all the tiles, in order to
allow straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extract contiguous cutouts of size 81.84\arcsec\ (440 pixels) with an
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts. This  results in a catalog of $\sim$ 430 000 cutouts for the entire \cfhtls
\textsc{wide} region. The size of the individual cutout is determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. It is possible that a lens
candidate happens to be too close to the edge of a cutout, the overlap
between neighbouring cutout allows the inspector to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, an inspector may not necessarily come
across the neighbouring cutout unless the inspector classifies a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications for the same cutout.

\subsection{Existing \cfhtls lens samples}
\label{sec:data:kls}

The \cfhtls data has been searched for lenses using semi-automated algorithms,
primarily, in the $g$ band where the predominantly  faint blue  source
galaxies are bright compared to the predominantly red deflector galaxies.
Here, we briefly mention the lens samples which were known to the authors
prior to the lens search with \sw.

At galaxy-scales, two primary lens searches have been carried out in
\cfhtls. The \rf \citep{Gavazzi2014} was used for finding compact rings
or arcs around centers of isolated and massive early-type galaxies. The
\rf subtracts the PSF-matched $i$-band images from the $g$-band images,
and looks for excess flux in the bluer $g$-band. An object detector
measures the properties of these residual blue features, and candidates
which meet the length-width ratio and tangential alignment criteria are
then visually inspected to form the final sample. \citet{Gavazzi2014}
pre-selected $\sim$638,000 targets as either photometrically-classified
early type galaxies, or objects selected to have red centers and blue
outer parts, from the T06 CFHTLS data release catalogs.  14370 were
found to show detectable blue residuals, and 2524 were visually
inspected, having passed the automatic feature selection process. This
led to a total of 330 lens candidates out of which 42 are good quality
(\texttt{q\_flag} = 3) and 288 are medium quality (\texttt{q\_flag} = 2)
candidates. In addition to this well defined sample, \citet{Gavazzi2014}
reported a further 71 serendipitously detected lens candidates.  From
this sample of ``\rf candidates,'' the SL2S team found, during their
follow-up campaign, 33 confirmed lenses (and 17 promising candidates).
{\bf AM: check if any of these are from \cfhtls DEEP?}

The second galaxy-scale lens search was to find edge-on galaxy lenses in the
\cfhtls \citep{Sygnet2010} by selecting galaxies based on profile fits in the
$i$-band performed using {\sc Sextractor} and with low inclination angle. This
sample has about 3 promising and a total of 18 lens candidates.
{\bf AM: perhaps, this should go in additional samples because this was not
used as one of the known samples,check ?}

Lastly, the \af \citep{More2012} was used for finding blue arc-like features in
the complete \cfhtls data without any pre-selection on the type of the lensing
object. The algorithm measures the second order moments of the flux distribution
in pixels within small regions to estimate the direction and extent of local
elongation of features. Pixels with high values of elongation are connected to
form an arc candidate. Finally, a set of thresholds on arc properties such as
the area, length, width, curvature and surface brightness are used to select
arc-like candidates. The search was carried out in the g-band which is the most
efficient wavelength to find typical lensed features.  This sample, called
SARCS, has 55 promising and a total of 127 lens candidates which are selected
from both \cfhtls WIDE and DEEP fields. The SARCS sample consists of some
galaxy-scale candidates and mostly groups/cluster scale lens candidates. This is
because more massive systems produce arcs or lensed images with large image
separation from the lensing galaxy which are easier to detect compared to the
galaxy-scales.  In the absence of a large systematically followed up
verification of the candidates, we use the  most promising sample of 26 lens
candidates as the bonafide known lens sample. And, the sample of 108 candidates
form the known lens candidates which come from the \cfhtls WIDE alone since the
\sw lens search is carried out on the WIDE imaging only.

For the purposes of transparency and to help with the training,
the volunteers participating in \sw-\cfhtls lens search were made aware of
these known lens samples. Images containing the systems from the above samples
were labelled as ``Known Lens Candidates'' in the Talk forum, where volunteers
have the opportunity to discuss their findings with fellow volunteers and the
science team. In this paper, we refer to the sample of 330 \rf and 108
\af lens candidates as the sample of ``known lens candidates'' and the sample of
confirmed (or most promising) 33 \rf and 26 \af as the sample of
``known lenses''.

\question{AM}{AV, add info on other lens samples from the literature?}

@@

We begin by considering all elliptical galaxies at $z<1$ in our parent \cfhtls
catalog as potential lens candidates for the simulated sample. To avoid using a
known lens galaxy for our simulation purpose, we exclude all those galaxies
whose positions match with the lensing galaxies from the known \cfhtls~SL2S lens
samples \cite{More2012} within 2~arcsec (XXX check).

@@

In order to calculate the lensing cross-section, we first calculate the
luminosity of each potential lensing galaxy using the photometric redshifts
($z_l$) from the LRG catalog. We use the $L-\sigma$ scaling relation from the
bright sample of \citep{Parker2007} to set the velocity dispersion of the halo
hosting the galaxy which will be later used in the model
\be
\label{magstar2}
\sigma_v=142 \left(\frac{L}{L_{*}}\right)^{1/3} \,.
\ee
We assume that the knee of the luminosity function of galaxies, $L_*$ evolves
such that there is a decline of $1.5$ magnitudes between $z=1$ to $z=0$
\citep{Faber2007}.\footnote{We anchor our $L_*$ evolution at low redshifts using the
determination of $L_*$ in the r-band by \citet{Blanton:2001}. To maintain
consistency in magnitude systems, we have converted the CFHT MegaCAM magnitudes
to SDSS magnitudes and k-corrected them to $z=0.1$.}

@@

Lens shear strength &  0.001 & 0.02 &  0.001 & 0.02 \\
Lens shear pa &  0 & 180 & 0 & 180  \\
Source ellipticity & 0.1 & 0.6 & & \\
Source PA & 0 & 180 & & \\
\hline
\end{tabular}
{ $\dagger$} -- corresponds to the factor used for Groups scale lenses.
\end{center}
\end{table}
\subsubsection{Groups-scale lenses}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{sw-cfhtls-figs/cand_fp_P_frac.pdf}
\caption{ \label{fig:stackP}
Distribution of different types of candidates as a function of the posterior
probability P. The types of the candidates are the FPs, the new candidates and
the known candidates. The new and known candidates have higher detection rate
for higher values of P, as expected.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\question{AM}{PJM,AV, 1. Any preference for which of the following words should be
used? My preference: volunteers (most of the times) and citizen scientists (some
times because it suits the context better)
2. Add anything else to this section?}
{\bf AM: Check and homogenize use of the words citizen scientists, citizens, users,
volunteers throughout the text.}

\section{Results}
\label{sec:results}

\subsection{\sw Lens Sample}

@@

We decide to inspect top few hundreds of subjects from a catalog of descending P
values as a reasonable sample size for visual inspection. We find 665 subjects
with P$>0.3$ in the catalog. These subjects are then visually inspected by three
lens experts (AM, AV, PJM). The subjects are assigned grades on a scale of 0 to
3, thus dividing the subjects into candidates that almost certainly do not
contain a lens, possibly a lens, probably a lens and almost certainly a lens,
respectively. We find XXX candidates with averaged grade $0<$G$<1.3$ (low
probability sample) and XXX candidates with G$\ge1.3$ (high probability sample).
Further information on the low probability sample such as their positions and
images are available at XXX. Among the high probability sample, there are XXX
known candidates and XXX new candidates. To avoid duplicity, further information
on the \sw detected known candidates sample is available at XXX whereas the new
lens candidate sample is presented below in \Sref{sec:results:newcand}.

>>>>

%%% ENSURE WORD LIMIT IS 250
We report the discovery of 28 promising and 58 total new lens candidates
in the CFHT Legacy Survey data (\cfhtls) based on about 11 million
classifications performed by citizen scientists as part of the first
blind lens search undertaken by \sw.  The principle goal of this lens
search was to identify lenses missed by lens finding robots (the \rf on
galaxy scales and \af on group/cluster scales), which have been
previously used to mine the \cfhtls data for lenses.  We compare the
properties of known lens samples detected by these algorithms to the \sw
sample and found that they are broadly similar.  The image separation
distribution calculated from the combined sample of known and new lens
candidates shows that our previous constraints on the average density
profile of the lens population are robust.  In addition to recovering
about 60\% of the known sample, the new \sw candidates possess a richer
variety compared to the lenses found by the two robots. We find that
analyzing only those classifications which are performed by the more
skilled volunteers, \sw can achieve a detection rate of up to 75\% for
the known lens sample.  Training and calibration of the performance of
citizen scientists is crucial for the success of \sw. We also present
the SIMCT pipeline, used for generating a sample of realistic simulated
lensed images in the CFHTLS, and a sample of duds and false positives
used in the training.  Such a training sample has a legacy value for
testing future lens finding algorithms. We make our training sample
publicly available.

\end{abstract}

@@

The last few decades have seen a rise in the discoveries of strong
gravitational lenses owing to the plethora of interesting applications
lenses have in astrophysics and cosmology. Strong lenses are routinely
used to probe the dark matter distribution from galaxy
\citep[e.g.][]{Koopmans2006,Barnabe2009,Sonnenfeld2015} to groups-cluster
scales
\citep[e.g.][]{Limousin2008,Zitrin2011,Oguri2012,More2012,Newman2013}, to
study distant young galaxies by using the lensing magnification as a
natural telescope \citep[e.g.][]{Zitrin2009,Zheng2012,Whitaker2014}, to
test the cosmological model by constraining cosmological parameters such
as the Hubble constant and dark energy
\citep[e.g.][]{Suyu2010,Collett2012,Collett2014,Sereno2014} and many more. Strong
lenses are rare since a foreground massive object needs to be
sufficiently aligned with a distant background source to produce
multiple images. Nevertheless, systematic lens searches have led to the
discovery of over 500 lenses till
date\footnote{http://masterlens.astro.utah.edu}.

The search for gravitational lenses is a needle-in-a-haystack problem.
Efficient automated methods are imperative to find a reasonably complete
and pure sample of strong lenses. Several lens finding algorithms have
been developed so far for this purpose
\citep[e.g.][]{Lenzen2004,Alard2006,Seidel2007,More2012,Gavazzi2014} but they can not
simultaneously capture the myriad types of lenses that are known to
exist. For example, the lensed images of background galaxies show
variety in their surface brightness distributions, colors, light
profiles, shapes, structures and angular image separations.  Moreover,
many lensed images appear similar to features found commonly in galaxies
(such as spiral arms) or to artifacts in astronomical images (scattered
light around stars).  Almost all lens finding algorithms find it
difficult to distinguish these from the real lenses and thus, suffer
from a high rate of false positives.  To minimize this problem,
algorithms are often restricted to detect a very narrow class of lens
systems. However, even after such restrictions, robotic lens searches
have to always rely on visual screening to arrive at a list of plausible
candidates.

Recognising patterns amidst an array of noisy pixelated images is one of the
strengths of human brains. Humans are also capable of dealing with
multi-tier complex web of questions before arriving at a conclusion which may
not be always possible to automate. The algorithm by which our brains process a
task is extremely malleable, self-learning and self-evolving. Therefore it has
a huge potential for the discovery of exotic objects which do not quite fit a
set criteria, but are still very likely to be objects of interest. The lens
finding algorithms are not yet advanced enough to produce better performance
than visual classifications. Given this, participation of a large community of
volunteers to help with the visual identification of lenses is extremely
beneficial for the lensing community. In particular, as we enter the era of
large area imaging surveys (spanning thousands of square degrees), it is
the perfect time to tap into the potential of citizen science.

\GZ was the first citizen science project in astronomy, and addressed the
problem of how to  classify large numbers of galaxies by their morphology
\citep{Lintott2008}.  From the early results based on morphological
classifications of galaxies, to several new unexpected and interesting
discoveries, such as that of green pea galaxies and Hanny's Voorwerp, \GZ
has been able to harness the potential of citizen scientists.  Since then
astronomy and non-astronomy projects have been launched under the citizen
science web portal Zooniverse (http://zooniverse.org). The task of finding
gravitational lenses is significantly challenging given the fact that these
lenses are rare. To add to the challenge, not many citizen scientists are
expected to be aware of the phenomenon of gravitational lensing, and the
resulting image configurations.  With these significant challenges at hand, we
designed the \sw project to enable the discovery of lenses through citizen
science.  In Paper I \citep{Marshall2015}, we describe the design of \sw and how the entire
system functions as a discovery service.

%
%\subsection{Paper I: Overview}
%In Paper I, we give details of the classification interface used by the
%volunteers. The regular images called test subjects are served randomly from the
%database. Some of the test subjects are labelled as training subjects because
%either these have simulated lensed images superposed around real galaxies or
%there are some false positives in the subjects.  The sample of training subjects
%is much smaller compared to the sample of test subjects. The training subjects
%are shown with decreasing frequency to the volunteers depending on how new the
%volunteers are to the project based on their classification history. However,
%the frequency of training subjects becomes a constant after XXX classifications
%from a volunteer.
%
%Clicking (or failing to click) on the training subjects results in a
%pop-up feedback message that congratulates the volunteers or points them
%to read the Spotter's guide pages. The website has several examples of
%real lenses and false positives with explanation on the Spotter's guide
%page. A quick reference to the Spotter's guide is available in the
%classification interface as well. Additional features to change zoom,
%contrast and brightness are made available to make a better judgement of
%the difficult lens candidates.
%
%The classification of subjects happens in two serial stages. \StageOne is
%for quick rejection of all uninteresting subjects. Subjects that passed
%\StageOne are then reinspected at \StageTwo, for a careful selection of
%most promising lens candidates. In general, subjects are assigned a
%small prior probability for it to contain a lens. As each subject is
%seen by multiple volunteers, the posterior probability P is updated by
%using the weights of the volunteers. The weights are determined based on
%how well the volunteers perform on the training sample. Once a subject
%crosses a threshold (acceptance or rejection), it is retired from the
%database to ensure that rest of the subjects receive sufficient multiple
%classifications. This is true for \StageOne with ~10$^5$ subjects but is
%not required for \StageTwo with only a few thousand subjects.
%
%\question{AM}{PJM,AV, should we have this overview section on paper I?}
%

In this paper ( Paper II hereafter), we describe our first lens search using
data from the Canada-France-Hawaii Telescope Legacy Survey
\citep[CFHTLS,][]{Gwyn2012}. This
paper is organised as follows. In \Sref{sec:data}, we introduce the \cfhtls
imaging data and the known lens samples from the \cfhtls. In \Sref{sec:ts} and
\Sref{sec:dfp}, we explain the training sample consisting of simulated lenses,
duds and false positives generated to aid the \sw volunteers in the lens search.
In \Sref{sec:results}, we present the new lens candidates from \sw and compare
\sw to two lens finding algorithms which produced the known \cfhtls lens
samples. Next, we discuss what kind of lenses are detected or missed by the
algorithms and \sw in \Sref{sec:discuss}. The summary and conclusions are given
in \Sref{sec:conclude}.

@@

The \cfhtls is a photometric survey in five optical bands
($u^*g'r'i'z'$) carried out with the wide-field imager MegaPrime which
has a 1~deg$^2$ field-of-view and a pixel size of 0.186\arcsec. The
\cfhtls \textsc{WIDE} covers a total non-overlapping area of 150~deg$^2$
on the sky and consists of four fields W1, W2, W3 and W4. The field W1
has the largest sky coverage of 63.65~deg$^2$. The fields W2 and W4 have
similar sky coverages of 20.32~deg$^2$ and 20.02~deg$^2$,
respectively\footnote{These numbers are estimated from
http://terapix.iap.fr/cplt/table\_syn\_T0006.html}.  The field W3 has a
sky coverage of 42.87~deg$^2$ and is more than twice as large as W2 and
W4.

The \cfhtls imaging is very homogeneous and has good image quality. Most of
the lensed arcs are much brighter in the $g$ band thus, deep imaging in
this band is desirable. The limiting magnitude is 25.47 for the $g$ band
which goes the deepest among all of the five bands. The mean seeing in
the $g$ band is 0.78\arcsec. The zero point to convert flux to AB
magnitude for all bands is 30. These characteristics make \cfhtls ideal
to do visual inspection for finding lenses.  We use the data from the
final T0007 release taken from the Terapix
website\footnote{{http://terapix.iap.fr/cplt/T0006-doc.pdf}}
for this work.

@@

\subsection{Previously published lens samples from the \cfhtls}
\label{sec:data:kls}

The \cfhtls has been searched for lenses using various lens finding
methods and algorithms. Here, we give a brief summary of previously published lens
samples in chronological order.

In the thesis dissertation of \citet{Thanjavur2009}, 9 promising and 2
low probability candidates are reported which were discovered
serendipitiously. These detections were made during the visual
inspection of the \cfhtls images as part of data reduction procedures
for the Weak Lensing survey \citep{Benjamin2007}.

\citet{Sygnet2010} carried out a search for edge-on galaxy lenses in the
\cfhtls WIDE. They identified galaxies, using {\sc Sextractor}, which
have $18<i<21$ and only those with inclination angle $<25$~deg. With a
few more selection criteria and visual inspection, they found about 3
promising and a total of 18 lens candidates.

The \af \citep{More2012} was used for finding blue arc-like features in
the complete \cfhtls data without any pre-selection on the type of the lensing
object. The algorithm measures the second order moments of the flux distribution
in pixels within small regions to estimate the direction and extent of local
elongation of features. Pixels with high values of elongation are connected to
form an arc candidate. Finally, a set of thresholds on arc properties such as
the area, length, width, curvature and surface brightness are used to select
arc-like candidates. The search was carried out in the g-band which is the most
efficient wavelength to find typical lensed features.  This sample, called
SARCS, has 55 promising and a total of 127 lens candidates which are selected
from both \cfhtls WIDE and DEEP fields. The SARCS sample consists of some
galaxy-scale candidates and mostly groups/cluster scale lens candidates. This is
because more massive systems produce arcs or lensed images with large image
separation from the lensing galaxy which are easier to detect compared to the
galaxy-scales.  In the absence of a large systematically followed up
verification of the candidates, we use the  most promising sample of 26 lens
candidates as the bonafide known lens sample. And, the sample of 108 candidates
form the known lens candidates which come from the \cfhtls WIDE alone since the
\sw lens search is carried out on the WIDE imaging only.

In \citet{Elyiv2013}, they visually inspected a sample of 5500 optical
counterparts of X-ray point-like sources identified in the XMM-LSS
imaging of the \cfhtls W1 field. The goal was to discover lensed
quasars. They find 3 very promising and a total of 18 candidates.

The \rf \citep{Gavazzi2014} was used for finding compact rings or arcs
around centers of isolated and massive early-type galaxies. The \rf
subtracts the PSF-matched $i$-band images from the $g$-band images, and
looks for excess flux in the bluer $g$-band. An object detector measures
the properties of these residual blue features, and candidates which
meet the length-width ratio and tangential alignment criteria are then
visually inspected to form the final sample. \citet{Gavazzi2014}
pre-selected $\sim$638,000 targets as either photometrically-classified
early type galaxies, or objects selected to have red centers and blue
outer parts, from the T06 CFHTLS data release catalogs.  14370 were
found to show detectable blue residuals, and 2524 were visually
inspected, having passed the automatic feature selection process. This
led to a total of 330 lens candidates out of which 42 are good quality
(\texttt{q\_flag} = 3) and 288 are medium quality (\texttt{q\_flag} = 2)
candidates. In addition to the main well-defined sample,
\citet{Gavazzi2014} reported a further 71 candidates were reported
detected by earlier version of the \rf or from the \cfhtls DEEP.  From
the main sample of ``\rf candidates,'' the SL2S team found, during their
follow-up campaign, 33 confirmed lenses (and 17 promising candidates).

The work by \citet{Maturi2014} used the arc finding code of
\citet{Seidel2007} and color properties of typical arcs to optimize arc
detection. This new approach was tested on the CFHTLS-Archive-Research
Survey \citep[CARS,][]{Erben2009} which covers an area of only 37~sq.~deg
and the entire data was visually inspected to estimate completeness and
purity of their robotic search. They found 29 candidates with the
algorithm and 41 candidates through visual inspection alone some of
which are known from previous searches. Most of these candidates are low
probability\footnote{http://www.ita.uni-heidelberg.de/˜maturi/Public/arcs}.

The \rf and the \af searches are the only searches that make use of a
lens finding algorithm that has been run on the entire \cfhtls imaging.
Thus, these can be considered as the known reference sample from robotic
searches.  For the purposes of transparency and to help with the
training, the volunteers participating in \sw-\cfhtls lens search were
made aware of these two known lens samples. Images containing the
systems from the \rf and the \af samples were labelled as ``Known Lens
Candidates'' in the Talk forum, where volunteers have the opportunity to
discuss their findings with fellow volunteers and the science team. In
this paper, we refer to the sample of 330 \rf and 108 \af lens
candidates as the sample of ``known lens candidates'' and the sample of
confirmed (or most promising) 33 \rf and 26 \af as the sample of ``known
lenses''. Note that the ``known lens'' sample is a subset of the ``known lens
candidates'' sample.

@@

In order to perform a blind lens search over the entire \cfhtls
\textsc{wide}, we present the volunteers with cutouts of images selected
from the survey region. We briefly describe the image presentation here
for completeness but more information can be found in Paper I. We use
the $g$, $r$ and $i$-band imaging from \cfhtls which are most useful for
visual identification of lenses.  We made color composite images using
the publicly-available code, HumVI\footnote{The open source  color image
composition code used in this work is available from
\texttt{http://github.com/drphilmarshall/HumVI}} following the
prescription of \citet{Lupton2004}. The color scales are chosen to
maximize the contrast between faint extended objects. These parameters
are then fixed during the production of all the tiles, in order to allow
straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extract contiguous cutouts of size 81.84\arcsec\ (440 pixels) with an
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts. This  results in a catalog of $\sim$ 430 000 cutouts for the entire \cfhtls
\textsc{wide} region. The size of the individual cutout is determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. It is possible that a lens
candidate happens to be too close to the edge of a cutout, the overlap
between neighbouring cutout allows the inspector to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, an inspector may not necessarily come
across the neighbouring cutout unless the inspector classifies a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications for the same cutout.
%We also provided access to the $u$ and $z$ band imaging
%data through dashboard, when the volunteers wished to take a closer look at a particular system\footnote{The dashboard tool page can be accessed from
%http://tools.zooniverse.org/\#/dashboards/spacewarp}.

%The code rescales the pixel values of each channel image into flux units, and
%then applies an arcsinh stretch. The stretch parameters are chosen
%using a small random sample of images, to ensure that the background
%noise is just visible, and that the centres of bright, intermediate
%redshift galaxies are not saturated.

@@

We begin by considering all elliptical galaxies at $z<1$ in our parent \cfhtls
catalog as potential lens candidates for the simulated sample. To avoid using a
known lens galaxy for our simulation purpose, we exclude all those galaxies
whose positions match with the lensing galaxies from the SARCS
samples within 2~arcsec (XXX check).

@@

In order to calculate the lensing cross-section, we first calculate the
luminosity of each potential lensing galaxy using the photometric redshifts
($z_l$) from the LRG catalog. We use the $L-\sigma$ scaling relation from the
bright sample of \citep{Parker2005} to set the velocity dispersion of the halo
hosting the galaxy which will be later used in the model
\be
\label{magstar2}
\sigma_v=142 \left(\frac{L}{L_{*}}\right)^{1/3} \,.
\ee
We assume that the knee of the luminosity function of galaxies, $L_*$ evolves
such that there is a decline of $1.5$ magnitudes between $z=1$ to $z=0$
\citep{Faber2007}.\footnote{We anchor our $L_*$ evolution at low redshifts using the
determination of $L_*$ in the r-band by \citet{Blanton:2001}. To maintain
consistency in magnitude systems, we have converted the CFHT MegaCAM magnitudes
to SDSS magnitudes and k-corrected them to $z=0.1$.}

@@

Lens shear strength &  0.001 & 0.02 &  0.001 & 0.02 \\
Lens shear pa &  0 & 180 & 0 & 180  \\
Source ellipticity & 0.1 & 0.6 & & \\
Source PA & 0 & 180 & & \\
\hline
\end{tabular}
{ $\dagger$} -- corresponds to the factor used for Groups scale lenses.
\end{center}
\end{table}
%{\bf check the units and confirm the numbers in this table etc.}

\subsubsection{Groups-scale lenses}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{sw-cfhtls-figs/cand_fp_P_frac.pdf}
\caption{ \label{fig:stackP}
Distribution of different types of candidates as a function of the posterior
probability P. The types of the candidates are the FPs, the new candidates and
the known candidates. The new and known candidates have higher detection rate
for higher values of P, as expected.}
\end{center}
\end{figure}

\section{Methodology to produce the \sw-\cfhtls Lens Sample}

@@

We decided to inspect the top few hundreds of images from a catalog of
descending P values as a reasonable sample size for visual inspection.
We find a total of 665 subjects with P$>0.3$ in the catalog. These
images were then visually inspected by three lens experts (AM, AV, PJM).
The images are assigned grades on a scale of 0 to 3, thus dividing the
subjects into candidates that almost certainly do not contain a lens,
possibly a lens, probably a lens and almost certainly a lens,
respectively. The final sample of \sw-\cfhtls lens candidates (see
\Sref{sec:swlens}) is produced by selecting candidates above some
threshold on the averaged grades.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}

\subsection{\sw-\cfhtls Lens Sample}
\label{sec:swlens}
